Comparing and combining network-guided biomarker discovery to discover genetic susceptibility mechanisms to breast cancer on the GENESIS study

### 2020-10-16 manuscript/plos/review_1.tex
address more comments

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./figures/figure_1.png}
  \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset. } As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it was excluded. Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level. \textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors. \textbf{(B)}~VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods. \textbf{(C)}~Solution networks produced by the different methods. \textbf{(D)}~Manhattan plots of SNPs/genes; in black, the method's solution. The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).}
  \label{fig:solution_overview}
\end{figure}

@@

Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results. We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}). Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise. We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once. The resulting stability-based consensus was composed of 68 genes (Fig \ref{fig:stable-consensus}). This network shares most of the properties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.

>>>>

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./figures/figure_1.png}
  \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset. } As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it was excluded. Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level. \textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors (Sections \ref{methods:algorithm_comparison} and \ref{methods:snp2gene}). \textbf{(B)}~VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods. \textbf{(C)}~Solution networks produced by the different methods. \textbf{(D)}~Manhattan plots of SNPs/genes; in black, the method's solution. The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).}
  \label{fig:solution_overview}
\end{figure}

@@

Most of the network methods, including the consensus, were highly unstable (Fig~\ref{fig:benchmark}), raising questions about the reliability of the results. We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}). Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise. We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once. The resulting stability-based consensus was composed of 68 genes (Fig \ref{fig:stable-consensus}). This network shares most of the properties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
